---
title: "Pokemon dual type analysis"
author: "Juan De Dios Santos, @jdiossantos"
output: html_document
---

```{r, echo=FALSE, include=FALSE}
require(ggplot2)
require(RColorBrewer)
require(tree)
require(tree)
set.seed(50)
setwd("~/Development/R/PokemonDualTypeAnalysis")
load("pokemontypesdf.RData")
types_matrix <- table(types)
```

# Introduction
***

In the world of Pokémon there are 18 types that are properties of the Pokémon and 
their moves. A Pokémon may have one or two types (dual type), meaning that there 
are 324 posible combinations, 18 single type Pokémon and 306 dual types Pokémon. 
However in the games, there are no differences between the order of the types in 
a dual type Pokémon, in other words, a Flying/Dragon Pokémon is treated in the 
same way a Dragon/Flying Pokémon is treated.

In this work we will take a look at dual type Pokémon. The analysis is mostly an
exploratory data analysis where we will study topics such as most frequent secondary
type by primary type, the most common secondary types by region and in the last
part, we will introduce a prediction model to see if there is any relation between
dual types Pokémon and their region of origin. 

Unlike the games, we will consider the order of the types of a dual type Pokémon
using the way it is documented in the Pokedex. For example, we will treat a 
Flying/Dragon Pokémon such as Noivern and Dragon/Flying Pokémon like Altaria as
having two different types. 

The Pokémon included in this study are the ones currently available in the National 
Pokedex, starting with Bulbasaur and ending with Zygarde; a total of 718 Pokémon. 
Also, Mega Pokémon and alternate forms of Pokémon like Rotom forms and the 
Therian forms are not included.

# The dataset
***

Our original dataset consists of two columns called `dataset.Type.1` and 
`dataset.Type.2`. The first column, `dataset.Type.1` has the primary type of the
Pokémon and the second one, `dataset.Type.2`, the secondary type. If the Pokémon
does not have a secondary type, the corresponding entry in `dataset.Type.2` is 
`<NA>` (not applicable). The dataset has a total of 718 observations (rows) 
because we are working with the National Pokedex and you may have noticed, the 
name of the Pokémon is not included, the reason behind this is that it is not 
really necessary since we are just interested in the types. Instead, we use the
index or the number of the row. For example, row number 4 is a pure Fire type 
Pokémon (Charmander).

```{r, echo=FALSE}
head(types)
```

#Overview of dual types

The following matrix shows the total of dual types Pokémon per each primary type. 
Each row represents a primary type and the columns the secondary type. For example,
the matrix shows there are 2 Bug/Electric Pokémon, 1 Bug/Fighting, 2 Bug/Fire 
and so on. The -1 entries means that it does not apply, since for example a Bug/Bug
Pokémon would be a pure Bug type, and pure types are out of the score of this analysis.

```{r, echo=FALSE}
diag(types_matrix) <- -1
types_table_complete <- as.data.frame(table(types))
types_table_complete$Freq <- as.numeric(types_table_complete$Freq)
# Frequency greater than 0
types_table <- types_table_complete[which(types_table_complete$Freq != 0), ]
types_table <- types_table[order(-types_table$Freq),]
types_matrix
```

The next figure, a heat map, shows the same information from the previous table 
in another way. The Y axis is the primary type and the X axis the secondary one.
Regardings the colors, the darker they are, the smaller the frequency is and the 
lighest the color is, higher the frequency is.

```{r, echo=FALSE}
color_palette <- colorRampPalette(rev(brewer.pal(6, "Greys")), space="rgb")
zp1 <- ggplot(as.data.frame(types_matrix),
              aes(x = dataset.Type.2, y = dataset.Type.1, fill = Freq))
zp1 <- zp1 + geom_tile()
zp1 <- zp1 + scale_fill_gradientn(colours = color_palette(50))
zp1 <- zp1 + coord_equal()
zp1 <- zp1 + theme_bw()
zp1 <- zp1 + theme(axis.text.x = element_text(angle=90, vjust=1))
print(zp1)
```


## Frequent secondary type by primary type

The figures in the section shows the quantity of secondary types by primary type. 
For example the first plot can be interpreted as this, "there are 13 Bug/Flying 
Pokémon, 11 Bug/Poison and so on."

### Bug dual types

```{r, echo=FALSE}
bug_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Bug", ]
# Remove rows where dataset.Type.1 and dataset.Type.2 are equal.
bug_dual_type <-
  bug_dual_type[bug_dual_type$dataset.Type.2 != bug_dual_type$dataset.Type.1, ] 
print(bug_dual_type[order(-bug_dual_type$Freq), ], row.names = FALSE)
bug_dual_type$ordered_Levels <- reorder(bug_dual_type$dataset.Type.2,
                                        -bug_dual_type$Freq)

ggplot(bug_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#A8B820") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  scale_y_continuous(breaks = seq(0,15)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Bug type Pokémon")
```


### Dark dual types

```{r, echo=FALSE}
dark_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Dark", ]
dark_dual_type <- 
  dark_dual_type[dark_dual_type$dataset.Type.2 != dark_dual_type$dataset.Type.1, ]
print(dark_dual_type[order(-dark_dual_type$Freq), ], row.names = FALSE)

dark_dual_type$ordered_Levels <- reorder(dark_dual_type$dataset.Type.2,
                                         -dark_dual_type$Freq)

ggplot(dark_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#705848") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Dark type Pokémon")
```

### Dragon dual types

```{r, echo=FALSE}
dragon_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Dragon", ]
dragon_dual_type <- 
  dragon_dual_type[dragon_dual_type$dataset.Type.2 != dragon_dual_type$dataset.Type.1, ]
print(dragon_dual_type[order(-dragon_dual_type$Freq), ], row.names = FALSE)

dragon_dual_type$ordered_Levels <- reorder(dragon_dual_type$dataset.Type.2,
                                           -dragon_dual_type$Freq)

ggplot(dragon_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#705898") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") + ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Dragon type Pokémon")
```

### Electric dual type

```{r, echo=FALSE}
electric_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Electric", ]
electric_dual_type <- 
  electric_dual_type[electric_dual_type$dataset.Type.2 != electric_dual_type$dataset.Type.1, ]
print(electric_dual_type[order(-electric_dual_type$Freq), ], row.names = FALSE)

electric_dual_type$ordered_Levels <- reorder(electric_dual_type$dataset.Type.2,
                                             -electric_dual_type$Freq)

ggplot(electric_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#F8D030") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) + 
  xlab("Secondary types") + ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Electric type Pokémon")
```

### Fairy dual type

```{r, echo=FALSE}
fairy_dual_type<-types_table_complete[types_table_complete$dataset.Type.1 == "Fairy", ]
fairy_dual_type <- 
  fairy_dual_type[fairy_dual_type$dataset.Type.2 != fairy_dual_type$dataset.Type.1, ]
print(fairy_dual_type[order(-fairy_dual_type$Freq), ], row.names = FALSE)

fairy_dual_type$ordered_Levels <- reorder(fairy_dual_type$dataset.Type.2,
                                          -fairy_dual_type$Freq)

ggplot(fairy_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#EE99AC") +
  scale_y_continuous(breaks=c(0,1,2)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Fairy type Pokémon")
```

### Fighting dual type

```{r, echo=FALSE}
fighting_dual_type<-types_table_complete[types_table_complete$dataset.Type.1 == "Fighting", ]
fighting_dual_type <-
  fighting_dual_type[fighting_dual_type$dataset.Type.2 != fighting_dual_type$dataset.Type.1, ]
print(fighting_dual_type[order(-fighting_dual_type$Freq), ], row.names = FALSE)

fighting_dual_type$ordered_Levels <- reorder(fighting_dual_type$dataset.Type.2, -fighting_dual_type$Freq)

ggplot(fighting_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#C03028") +
  scale_y_continuous(breaks=c(0,1,2)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Fighting type Pokémon")
```

### Fire dual type

```{r, echo=FALSE}
fire_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Fire", ]
fire_dual_type <-
  fire_dual_type[fire_dual_type$dataset.Type.2 != fire_dual_type$dataset.Type.1, ]
print(fire_dual_type[order(-fire_dual_type$Freq), ], row.names = FALSE)

fire_dual_type$ordered_Levels <- reorder(fire_dual_type$dataset.Type.2,
                                         -fire_dual_type$Freq)

ggplot(fire_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#F08030") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Fire type Pokémon")
```

### Flying dual types

```{r, echo=FALSE}
flying_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Flying", ]
flying_dual_type <-
  flying_dual_type[flying_dual_type$dataset.Type.2 != flying_dual_type$dataset.Type.1,]
print(flying_dual_type[order(-flying_dual_type$Freq), ], row.names = FALSE)

flying_dual_type$ordered_Levels <- reorder(flying_dual_type$dataset.Type.2,
                                           -flying_dual_type$Freq)

ggplot(flying_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#A890F0") +
  scale_y_continuous(breaks = c(0,1,2)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Flying type Pokémon")
```

### Ghost dual types

```{r, echo=FALSE}
ghost_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Ghost", ]
ghost_dual_type <-
  ghost_dual_type[ghost_dual_type$dataset.Type.2 != ghost_dual_type$dataset.Type.1,]
print(ghost_dual_type[order(-ghost_dual_type$Freq), ], row.names = FALSE)

ghost_dual_type$ordered_Levels <- reorder(ghost_dual_type$dataset.Type.2,
                                          -ghost_dual_type$Freq)

ggplot(ghost_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#705898") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Ghost type Pokémon")
```

### Grass dual types

```{r, echo=FALSE}
grass_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Grass", ]
grass_dual_type <- 
  grass_dual_type[grass_dual_type$dataset.Type.2 != grass_dual_type$dataset.Type.1,]

print(grass_dual_type[order(-grass_dual_type$Freq), ], row.names = FALSE)

grass_dual_type$ordered_Levels <- reorder(grass_dual_type$dataset.Type.2,
                                          -grass_dual_type$Freq)

ggplot(grass_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#78C850") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Grass type Pokémon")
```

### Ground dual types

```{r, echo=FALSE}
ground_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Ground", ]
ground_dual_type <- 
  ground_dual_type[ground_dual_type$dataset.Type.2 != ground_dual_type$dataset.Type.1, ]
print(ground_dual_type[order(-ground_dual_type$Freq),], row.names = FALSE)

ground_dual_type$ordered_Levels <- reorder(ground_dual_type$dataset.Type.2,
                                           -ground_dual_type$Freq)

ggplot(ground_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#E0C068") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Ground type Pokémon")
```

### Ice dual types

```{r, echo=FALSE}
ice_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Ice", ]
ice_dual_type <- ice_dual_type[ice_dual_type$dataset.Type.2 != ice_dual_type$dataset.Type.1, ]
print(ice_dual_type[order(-ice_dual_type$Freq),], row.names = FALSE)

ice_dual_type$ordered_Levels <- reorder(ice_dual_type$dataset.Type.2, -ice_dual_type$Freq)

ggplot(ice_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#98D8D8") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Ice type Pokémon")
```

### Normal dual types

```{r, echo=FALSE}
normal_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Normal", ]
normal_dual_type <- 
  normal_dual_type[normal_dual_type$dataset.Type.2 != normal_dual_type$dataset.Type.1,]
print(normal_dual_type[order(-normal_dual_type$Freq),], row.names = FALSE)

normal_dual_type$ordered_Levels <- reorder(normal_dual_type$dataset.Type.2,
                                           -normal_dual_type$Freq)

ggplot(normal_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#A8A878") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Normal type Pokémon")
```

### Poison dual types

```{r, echo=FALSE}
poison_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Poison", ]
poison_dual_type <- 
  poison_dual_type[poison_dual_type$dataset.Type.2 != poison_dual_type$dataset.Type.1, ]
print(poison_dual_type[order(-poison_dual_type$Freq), ], row.names = FALSE)

poison_dual_type$ordered_Levels <- reorder(poison_dual_type$dataset.Type.2,
                                           -poison_dual_type$Freq)

ggplot(poison_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#A040A0") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Poison type Pokémon")
```

### Psychic dual type

```{r, echo=FALSE}
psychic_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Psychic", ]
psychic_dual_type <- 
  psychic_dual_type[psychic_dual_type$dataset.Type.2 != psychic_dual_type$dataset.Type.1, ]
print(psychic_dual_type[order(-psychic_dual_type$Freq), ], row.names = FALSE)

psychic_dual_type$ordered_Levels <- 
  reorder(psychic_dual_type$dataset.Type.2, -psychic_dual_type$Freq)

ggplot(psychic_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#F85888") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") + ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Psychic type Pokémon")
```

### Rock dual type

```{r, echo=FALSE}
rock_dual_type <- types_table_complete[types_table_complete$dataset.Type.1 == "Rock", ]
rock_dual_type <- rock_dual_type[rock_dual_type$dataset.Type.2 != rock_dual_type$dataset.Type.1, ]
print(rock_dual_type[order(-rock_dual_type$Freq), ], row.names = FALSE)

rock_dual_type$ordered_Levels <- reorder(rock_dual_type$dataset.Type.2, -rock_dual_type$Freq)

ggplot(rock_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#B8A038") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") +
  ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Rock type Pokémon")
```

### Steel dual type

```{r, echo=FALSE}
steel_dual_type<-types_table_complete[types_table_complete$dataset.Type.1 == "Steel", ]
print(steel_dual_type[order(-steel_dual_type$Freq), ], row.names = FALSE)

steel_dual_type <- 
  steel_dual_type[steel_dual_type$dataset.Type.2 != steel_dual_type$dataset.Type.1, ]
steel_dual_type$ordered_Levels <- 
  reorder(steel_dual_type$dataset.Type.2, -steel_dual_type$Freq)

ggplot(steel_dual_type, aes(x=(ordered_Levels), y = Freq)) +
  geom_bar(binwidth=1, stat="identity", fill="#B8B8D0") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") + ylab("Frequency of secondary types") +
  ggtitle("Frequency of secondary types on Steel type Pokémon")
```

### Water dual type

```{r, echo=FALSE}
water_dual_type<-types_table_complete[types_table_complete$dataset.Type.1 == "Water", ]
print(water_dual_type[order(-water_dual_type$Freq), ], row.names = FALSE)

water_dual_type <- 
  water_dual_type[water_dual_type$dataset.Type.2 != water_dual_type$dataset.Type.1, ]

water_dual_type$ordered_Levels <- 
  reorder(water_dual_type$dataset.Type.2, -water_dual_type$Freq)

ggplot(water_dual_type, aes(x=(ordered_Levels), y = Freq)) + 
  geom_bar(binwidth=1, stat="identity", fill="#6890F0") + 
  scale_y_continuous(breaks = c(0,2,4,6,8,10)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1)) +
  xlab("Secondary types") + ylab("Frequency of secondary types") + 
  ggtitle("Frequency of secondary types on Water type Pokémon")
```

# Frequency of secondary types
***

This table shows the frequency of secondary types.
For example, there are 3 Pokémon with Bug as a secondary type (Anorith, Armaldo
and Skorupi).

```{r, echo=FALSE}
types_df <- as.data.frame.matrix(types_matrix)
diag(types_df) <- 0
colSums(types_df)
```

The most common secondary type is Flying with 87 and the rarest is Bug with just 3.

# Secondary types by region
***

In this section we will summarize the dual types per region (Kanto, Johto, Hoenn,
Sinnoh, Unova and Kalos).
The following tables shows the top five dual types per region.

```{r, echo=FALSE}
# Top 10 per region
region <- vector(length=718)
region[1:151] <- "Kanto"
region[152:251] <- "Johto"
region[252:386] <- "Hoenn"
region[387:493] <- "Sinnoh"
region[494:649] <- "Unova"
region[650:718] <- "Kalos"

types$region <- region  # Add the column to the types matrix.
types <- types[complete.cases(types),]  # Remove NA
diag(types_matrix) <- -1
types_table_complete <- as.data.frame(table(types))
types_table_complete$Freq <- as.numeric(types_table_complete$Freq)

# Get those Pokemon with frequency greater than 0
types_table <- types_table_complete[which(types_table_complete$Freq != 0),] 
types_table <- types_table[order(-types_table$Freq),]
```

### Kanto region

```{r, echo=FALSE}
# Kanto
kanto_dual_types <- types_table[which(types_table$region == "Kanto"),]
print(head(kanto_dual_types), row.names = FALSE)
```

### Johto region

```{r, echo=FALSE}
# Johto
johto_dual_types <- types_table[which(types_table$region == "Johto"),]
print(head(johto_dual_types), row.names = FALSE)
```

### Hoenn region

```{r, echo=FALSE}
# Hoenn
hoenn_dual_types <- types_table[which(types_table$region == "Hoenn"),]
print(head(hoenn_dual_types), row.names = FALSE)
```

### Sinnoh region

```{r, echo=FALSE}
# Sinnoh
sinnoh_dual_types <- types_table[which(types_table$region == "Sinnoh"),]
print(head(sinnoh_dual_types), row.names = FALSE)
```

### Unova region

```{r, echo=FALSE}
# Unova
unova_dual_types <- types_table[which(types_table$region == "Unova"),]
print(head(unova_dual_types), row.names = FALSE)
```

### Kalos region

```{r, echo=FALSE}
# Kalos
kalos_dual_types <- types_table[which(types_table$region == "Kalos"),]
print(head(kalos_dual_types), row.names = FALSE)
```

```{r, echo=FALSE}
dual_types_freq_by_region <- data.frame(Region = character(), Frequency = numeric(), stringsAsFactors = FALSE)

dual_types_freq_by_region[1,] <- c("Kanto", as.numeric(sum(kanto_dual_types$Freq)))
dual_types_freq_by_region[2,] <- c("Johto", as.numeric(sum(johto_dual_types$Freq)))
dual_types_freq_by_region[3,] <- c("Hoenn", as.numeric(sum(hoenn_dual_types$Freq)))
dual_types_freq_by_region[4,] <- c("Sinnoh", as.numeric(sum(sinnoh_dual_types$Freq)))
dual_types_freq_by_region[5,] <- c("Unova", as.numeric(sum(unova_dual_types$Freq)))
dual_types_freq_by_region[6,] <- c("Kalos", as.numeric(sum(kalos_dual_types$Freq)))

dual_types_freq_by_region$ordered_levels <- reorder(dual_types_freq_by_region$Region,
                                                    -as.numeric(dual_types_freq_by_region$Frequency))

ggplot(dual_types_freq_by_region,aes(x=ordered_levels, y = Frequency)) + geom_bar(binwidth=1, stat="identity", fill="darkorange") + xlab("Region") + ggtitle("Frequency of dual types Pokemon across regions") 
```

The previous figure shows that the Unova region has the most dual type Pokémon and
Kalos has the least.

The next and last figure of this section shows the frequency of dual type Pokémon
per each unique type splitted by generation. The x axis shows the different types,
the y axis the frequency and each colored line represents a region. For example,
there are 15 Bug Pokémon with a secondary type originally from Unova.

```{r, echo=FALSE}
types_per_region <- as.data.frame(table(types$dataset.Type.1, types$region))
colnames(types_per_region) <- c("Type", "Region", "Frequency")
ggplot(types_per_region, aes(x=Type, y=Frequency, group=Region)) + geom_line(aes(colour=Region), size = 1) + geom_point(size = 3) + scale_colour_brewer(palette="Dark2") +
  scale_y_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15)) + theme(axis.text.x = element_text(angle = 90, vjust = 1)) + ggtitle("Frequency of Pokemon with dual types per primary type across generations") + theme(plot.title = element_text(size = rel(0.9)))
```

# Is there any relation between dual type Pokémon and their region of origin?
***
This section is in some way different than the previous ones. Here we will
introduce a classification model called a decision tree. A decision tree is a 
prediction model that uses a flow-chart structure to reach certain conclusions 
about a dataset based on previous observations. This simple decision tree was 
made to check if there is any relation between dual types Pokémon
and their region of origin (Kanto, Johto, Hoenn, Sinnoh, Unova and Kalos). 
For example, are most Normal-Flying Pokémon from the Kanto region? 

What we did in this decision tree model was to split the dataset into a training
set and a test set. The training set was used to teach the model about the dual types
Pokémon and their original region and the test set for detecting how well our model
can "guess" the original region of a certain dual type Pokémon based on what it learned.

Note that this model is not optimized at all. The purpose of doing this is just 
to see if there is any correlation at all.

```{r, echo=FALSE}
# Prepare the dataset
types$region <- as.factor(types$region)

# Train the decision tree
train <- sample(1:nrow(types), 200) #147 for test
tree.types <- tree(region ~ . , types, subset = train) 
summary(tree.types)

# Predict using those observations not seen during training
tree.pred <- predict(tree.types,types[-train,],type="class")
with(types[-train,],table(tree.pred,region))
```

If we count the correct predictions (the numbers in the diagonal) and using length
of the test set (147) we get a correct prediction percentage of 
`r (sum(diag(with(types[-train,],table(tree.pred,region))))/sum(with(types[-train,],table(tree.pred,region)))) * 100` which means that 1 out of 3 dual types Pokémon gets a correct 
prediction of its region based on its dual type. A 33% correct rate does not say
much about this model; as seen from the table there is some connection in Kanto 
dual type Pokémon and Unova but generally speaking my interpretation is that 
there is not a constant connection between a dual type Pokémon and its origin region.


The next figure is a representation of the tree. Each internal node represents a
test on the attribute, which in this case is the primary type and the secondary 
type of the Pokémon. For example, at the first internal node we have all the 
secondary types except for Electric, Fire, Poison and Steel. At this node we ask,
is our Pokémon secondary type any of these? If the answer if yes, we take the left
route otherwise the right now. Suppose the answer to this question is no, so we 
take the right branch. At this new branch we ask, is our Pokémon secondary type 
either Electric, Fire or Steel? If the answer is yes we take the left branch and
performed another test, otherwise we take the right split and the Pokémon is 
categorized as Kanto. In this exact case, the only remaining secondary type not 
seen in the branches is Poison, thus according to this tree, a Pokémon that has 
Poison as a secondary type comes from the Kanto region. In some sense this is 
correct because out of all the 31 Pokémon that has secondary Poison type, 19 comes
from the Kanto region (61.3%), 3 from Johto (9.7%), 2 from Hoenn (6.5%), 2 from 
Sinnoh (6.5%) and 5 from Unova (16%).

```{r, echo=FALSE, fig.width=10, fig.height=7}
plot(tree.types)
text(tree.types, cex = 0.6, pretty = 1)
```

# Notes and disclaimers
***

* If you spot a typo, or any error in the analysis, feel free to notify me.
* This work is licensed under a [Creative Commons Attribution-NonCommercial 4.0 International License](http://creativecommons.org/licenses/by-nc/4.0/). In other words, feel 
free to use it, share it, edit it for non-commercial purposes and please, give credit.
* ? 2015 Pokémon. ? 1995-2015 Nintendo/Creatures Inc./GAME FREAK inc. Pokémon, Pokémon character names are trademarks of Nintendo.
